### 2018-02-14 R/diagrams/heuristic.Rmd
Added markdown with heuristic documentation

---
title: "heuristic1"
author: "Franco Peschiera"
date: "January 9, 2018"
output:
    html_document:
        # css: ./../../css/baobab.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Basic idea

The algorithm consists on a greedy logic to fill, one mission at a time, with aircraft until the needed amount is reached. It tries to use each aircraft as much as possible until it reaches its maximum flight time. When this happens: the aircraft is granted a maintenance, if possible and as soon as possible. When the aircraft has all its slots full or can't be assigned a maintenance and needs one, the algorithm continues with the next aircraft.

## Pseudocode

With the candidate and the task we do the following:

    for every task:
        1. calculate candidates.
        for each candidate until we run out of candidates or we satisfy task's needs:
            1. find available periods.
            Until we run out of periods:
                1. we check if we can assign it (checking remaining time and elapsed time). 
                    * If we can't: we tag the aircraft as "in need of a maintenance'.
                    * if we can: we assign the task.
                2. After assigning a task: we update the remaining and elapsed times for the aircraft.

For assigning maintenances is more or less similar:

    if the aircraft is tagged as 'in need of a maintenance':
        1. we check if we can reserve 6 continous months of maintenance.
            * if we can: we do it (and continue assigining tasks to the aircraft).
            * if we cannot: we go to the next candidate.
        2. After assigning a maintenance: we update the remaining and elapsed times for the aircraft.

<!-- <br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br> -->
## Main python code

I think the code gives an idea of the logic behind it. I only include the code for each mission so we would have to execute this function for each mission (or task).

`task` is the mission to satisfy.

* **candidates**: a list with the universe of aircraft that can be assigned to that task.
* **rem_resources**: remaining aircraft that need to be assigned at each period of the task's duration.
* **periods_task**: intersection of available slots for the aircraft (periods without taks or maintenance) and the task's period's needs.

```python

def fill_mission(self, task):
    dtype = 'U7'
    # get task candidates: we copy the list.
    candidates = self.instance.data['tasks'][task]['candidates'][:]
    # get task periods to satisfy:
    rem_resources = \
        aux.dicttup_to_dictdict(
            self.instance.get_task_period_needs()
        )[task]

    while len(rem_resources) > 0 and len(candidates) > 0:
        # candidate = 'A124'
        candidate = candidates[0]
        # get free periods for candidate
        periods_task = \
            self.get_free_starts(candidate, np.fromiter(rem_resources, dtype=dtype))

        # consider eliminating the resource from the list.
        # if all its periods are 'used'
        if len(periods_task) == 0:
            candidates.pop(0)
            continue
        print('candidate={}\ntask={}\nperiods={}'.format(candidate, task, periods_task))
        maint_period = ''
        # TODO: this is still wrong:
        for start, end in periods_task:
            for period in aux.get_months(start, end):
                result = self.check_set_task(candidate, period, task)
                if not result:
                    maint_period = period
                    break
                rem_resources[period] -= 1
            self.update_time_usage(candidate, period, self.get_next_maintenance(candidate, period), 'rut')

        if maint_period != '':
            # find soonest period to start maintenance:
            maint_start = self.get_soonest_maint(candidate, maint_period)
            if maint_start is None:
                # this means that we cannot assign tasks BUT
                # we cannot assign maintenance either :/
                # we should then take out the candidate:
                candidates.pop(0)
                continue
            maint_end = aux.shift_month(maint_start, 5)
            periods_maint = aux.get_months(maint_start,  maint_end)
            for period in periods_maint:
                self.set_maint(candidate, period)
            self.update_time_maint(maint_start, maint_end)
            next_month = aux.get_next_month(maint_end)
            self.update_time_usage(candidate, next_month, 'ret')
            self.update_time_usage(candidate, next_month, 'rut')

```