### 2018-03-05 R/OptiPlan_presentation/proposal.Rmd
updated proposal with planning. Added french version. slighlty changed css

---
title: "Optiplan proposal"
date: "March 11, 2018"
output: 
  html_document:
    css: ./../../css/baobab_docs.css
    toc: true
  beamer_presentation:
    theme: "Berlin"
    toc: true
    slide_level: 3
    incremental: yes
    fig_height: 4
    # footer: "ISAE-SUPAERO, Université de Toulouse, France"
    includes:
        in_header: header.tex
  slidy_presentation:
    css: ./../../css/baobab_MH.css
    incremental: yes
    footer: "ISAE-SUPAERO, Université de Toulouse, France"
    slide_level: 4
    toc: true
nocite: '@*'
bibliography: MFMP.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
setwd("../")
source('scripts/params.R')
source('scripts/import_results.R')
source('diagrams/cases.R')
source("diagrams/architecture.R")
source("diagrams/functionalities.R")
source('diagrams/exampleDT.R')
exp_directory = PATHS[['experiments']] %>% paste0('201802141539/')
gantt <- print_solution(exp_directory, max_resources = 5, width="100%", height="200px")
graphs_rem <- graph_remaining(exp_directory)
table_rem <- table_remaining(exp_directory)
html_output <- knitr::opts_knit$get("rmarkdown.pandoc.to") %in% c("slidy", 'html')
```

<!-- TODO: check if all points from the rules are covered -->

## Introduction

In this document we present the problem as the team understands it. The information used was the previous knowledge the team and the OptiPlan meeting in Paris.

Following, a technical proposal is presented that includes the functionalities a possible complete application would include based on the exchanges made during the OptiPlan day in Paris and from previous experiences the team has in building planning applications for industry.

Furthermore, the team's credentials are presented, mentioning the different projects and research that is relevant to the present project.

Finally, the technologies and tools to be used are presented, describing the use and justifying the choice at each point.

## Summary

The project presented is a web-based modern interactive application. It uses a mathematical model to generate new solutions, permits different users to access it with different access rights and provides powerful visualization to show, compare, and analyze plannings. It integrates with existing data sources (Excel file), new proposed ones (Excel template) and possible existing ones (other internal systems) as well as simple output (csv, json). Finally, it permits the use of cases to store information, making it easy to test, experiment and compare different scenarios.

We believe this team is uniquely well-positioned to tackle this project because of the following reasons:

* The team has more than 40 years of experience in the Operations Research field, which is the theory behind this problem.
* The team has years of experience in building planning and optimization software applications for a wide range of big industrial companies.
* The team has vast experience in both scientific articles and commercial projects related to planning and maintenance.
* The team has unique knowledge of the specific problem at hand, because a PhD thesis with this subject was started in October 2017 in collaboration with the DGA. Work was presented in ROADEF 2018 and is going to be presented in MOSIM 2018.

## Problem

The problem is considered a Flight and Maintenance Planning problem. Work has been already done by @Kozanidis2008, @Cho2011, @Verhoeff2015 and @Chastellux2016, among others.

The problem consists in assigning aircraft to predefined missions and scheduling periodic preventive maintenance for these same aircraft.

Missions can only be understood as a budget of hours per month or per year. In this document we present the most general formulation possible, where missions are assigned to aircraft for one or more months.

### Missions

There is a fixed set of $j \in \mathcal{J}$ missions to be accomplished over an horizon of time divided into $t \in \mathcal{T}$ discrete periods. For their execution, these missions require the assignment of a specific number of aircraft $R_j$ each period of time the mission is active. The start and end periods for each mission are known and a mission is considered active between its start and end period.

During each period, missions consume an amount of time equal to $H_j$ hours from each of its assigned aircraft.

The assignment of an aircraft to a mission is not decided for the whole duration of the mission. After a minimum amount of time $MT_j$, an aircraft can be freed and exchanged for another one, even if the mission it is assigned to has not finished. 
The total number of aircraft being used at any given time in a specific mission should always be equal to $R_j$.

Each mission requires one and only one type of aircraft which, in addition, should comply with additional mission requirements.

### Aircraft

There is a set $i \in \mathcal{I}$ of available aircraft that are assigned to missions in order to accomplish them. Each aircraft can only be assigned to a single mission in any given period. These aircraft suffer from wear and tear and require regular maintenance operations during their lifetime. The need for maintenance is calculated based on two indicators.

The first one is called "remaining elapsed time" (or $ret_{it}$). It expresses the amount of time (measured in time periods) after which the aircraft cannot be used anymore and has to undergo a maintenance operation. Its value is calculated for each aircraft $i$ and each time period $t$. In a similar way, "remaining usage time" (or $rut_{it}$) is used to measure the amount of time that the aircraft $i$ can be used before needing a maintenance operation at any given period $t$.

Additionally, after an absolute amount of time and/or usage ($aet_i$ or $aut_i$), the aircraft becomes obsolete. There is no way to reverse this process.

At any given period, including at the start of the planning horizon, each aircraft has a specific status given by remaining usage time and remaining elapsed time.

### Maintenances

Maintenance operations are the process by which aircraft that have reached a limit in some indicator can return to a state where they can continue to be used in missions.

Each maintenance operation has a fix duration of $M$ periods.

After a maintenance operation, an aircraft restores its remaining elapsed time and remaining usage time to their max values $E$ (~60 months) and $H$ (~1200 hours) respectively.

There are also special maintenance operations that improve the aircraft capabilities (their ability to be assigned to new missions). These operations will be decided by the user and fixed for the application so they can be taken into account when assigning the rest of maintenance operations.

### Storages

Following the rule of remaining elapsed time, even if an aircraft is not being used, it still needs to have a maintenance after a given amount of time has passed. In order to avoid this problem, the aircraft can be put into a storage state.

A aircraft in this state has to be kept for a minimum time of $sm$ periods. While in this state it cannot receive maintenance or be assigned to any mission.

Every aircraft has the capacity to be stored and this capacity is measured in a number of periods $sc$. In order for an aircraft to restore its remaining storage capacity, it needs to receive a specific maintenance. Similar to the remaining elapsed time, the aircraft recovers its storage capacity up to a certain level $S$.

### Possible states

As a summary, the following are the possible logical states that an aircraft can be in:

* Assigned to a mission.
* Receiving maintenance.
* Under storage.
* Available.

### Time

In planning missions and maintenance operations, it is important to take into account the initial and end state of each aircraft. This initial state can be a maintenance or an assigned mission. If an aircraft is already under maintenance, it needs to continue in this state for its remaining maintenance time. Tasks' assignments should be taken into account in a similar manner.

For planned maintenance operations whose starting time is close to the beginning of the planning horizon, they should also be considered fixed.

The remaining used and elapsed times are assigned to each aircraft at the beginning of the planning horizon.

Finally, the state of each aircraft at the end of the planning horizon, its remaining (elapsed, used) time, needs to be defined and taken into account.

### Objectives

The following objectives are considered, although more could be potentially added.

Given that the creation of new missions and the duration of maintenance are considered stochastic in real-life, one basic goal is to maximize the robustness of the planning by having the greatest amount of available aircraft at every period of the planning horizon.

Given the limited amount of maintenance capacity and its cost, another goal is to smooth as much as possible the number of aircraft under maintenance over the planning horizon. Maintenance operations should also be controlled so it does not pass the maximum capacity.

## Proposal

A summary of the application's functionalities is shown in the following image:

```{r, echo=FALSE, out.width='70%'}
if (html_output){
    graph_functionalities()
} else {
    knitr::include_graphics('../../img/functionalities.png')
}
```

    Graph showing functionalities for the application

### Optimizer

The model built to generate new solutions will comply with the requirements described below.

#### RQ001: missions hours

Each month, all missions need to be satisfied by assigning enough flight hours to each aircraft.

#### RQ002: assigned aircraft

Each month, all missions need to have a minimum number of aircraft assigned.

#### RQ003: maintenance needs

Each aircraft needs to always have enough flight hours to be able to do the required missions it is assigned to. In order to recover its flight hours (to around 1200 hours), a maintenance operation is needed.

Similarly, after certain amount of time has passed, aircraft need a maintenance even if they did not reach this flight hour limit. This amount of time is around 5 years.

#### RQ004: maintenance duration

Maintenance operations take around 5 months and once an aircraft has started the maintenance operation, it needs to continue it. It cannot do any operation while in maintenance.

#### RQ005: availability

At each month, it is important to have the biggest possible number of available aircraft ready to engage in a new mission.

#### RQ006: maintenance capacity

The maintenance operations to be planned for all aircraft need to be done in the most smooth way possible, avoiding specific months were there are many aircraft under maintenance and other months were there are but a few aircraft under maintenance.

In other words, it is important to have the lowest possible number of available aircraft under maintenance at each month.

#### RQ007: end state

At the end of the planning horizon, the aircraft need to be in the best state possible. This implies that, in average, they need to have a big budget of flight hours remaining and a big number of months before needing a maintenance.

#### RQ008: storage

The storage can be decided by the model in order to avoid the deterioration of the aircraft. The minimum period to store an aircraft is 6 months and while in this state it is unavailable both for maintenance and for missions.

#### RQ009: fixing parts of the planning

There are several reasons for the user to want to fix some part of a previous planning. For example, there are some maintenance operations that are already decided and need to take place in the future. There are some aircraft that are already in a mission and need to stay there for some months at least. Also, storage can be already decided or in process.

Another reason for fixing maintenance operations could be for planned improvement operations to aircraft. These special type of maintenance operations are decided many months in advance for specific selected aircraft.

These fixed assignments will be communicated to the planning generator in order for it to not change them.

#### RQ010: configuration

Available options will be, by default:

* **Maximal time**: in case a solution is needed before proving optimality.
* **Maximal GAP**: in order to apply a maximum allowed distance from the best obtained solution to the hypothetical best possible solution. This distance can be a proportional difference or an absolute difference.
* **First and last month**: in order to define the horizon of planning to plan.
* **Objective weights**: for the different objectives to optimize, to give preference to one or another objective.
* **Maintenance parameters**: maintenance duration (in months), maximum flight hours, maximum elapsed time, and other maintenance-related parameters.

### Cases and experimentation

Given the strategic nature of the application, it will be case-centric. This means that it will support saving all the relevant data (input and output) in a case. This way, different experiments could be done by copying and modifying an existing case. This is similar as working with multiple spreadsheets, each one with its own name, while sharing the same logic in one place.

The following figure shows an hypothetical set of cases already saved:

The following figure represents the possible interactions between cases. Case 1 was published for the month of March. During the month, it received some modifications, resulting in cases 2 and 3. Finally, Case 3 was (re)published for March.
Later, Case 4 was created and modified before publication to Case 5. Case 5 was published for April.

```{r, echo=FALSE, out.width='70%'}
if (html_output){
    graph_cases()
} else {
    knitr::include_graphics('../../img/cases.png')
}
```

    Example showing interaction between cases and publication of cases.

The available functionality will be the following:

1. Create a new empty case.
2. Copy an existing case.
3. Delete a case.
4. Compare two cases.
5. Publish a case.
6. Export a case.

#### Contents of a case

A case consists on all the information needed to construct a planning and one possible planning built from this information. This unit of storage permits to easily integrate with current data structures and run, compare different experiments.

More explicitly:

* Set of aircraft with their characteristics and historic information on them until a certain date.
* Set of missions or flight hours to do in the planning period.
* Assignments of aircraft to operations during the planning horizon.
* Future maintenance schedules.
* Configuration such as maintenance duration, flight hours after a maintenance, etc.

#### Comparison of cases

This functionality will potentially work in several levels depending on the final needs of the client.
One option would be to visualize the differences in input data for two cases.
Another option would be to visualize the differences in the resulting planning for the two cases.
Finally, it could be interesting to compare specific KPI for two or more cases, in order to choose the best scenario.

#### Published case

A situation that is common in the creation of plannings is the fact that several different cases may exist but it is still important to identify which one is the 'active' one. In order to deal with this, a simple publication system will be included in the application, where the 'planner' (see *Users* section) can identify the case that should be shared with others.

### Graphic User Interface (GUI)

All interaction from the user with the application will be done via an HTML application.

Besides showing data in interactive tables, specific alternatives can be incorporated to make the usage easier and more potent.

Below are some examples of available functionality.

Here, an interactive gantt is used to present and potentially modify a maintenance plan. It is possible to edit the maintenance operations and assignments tasks (change the size or move it) by double clicking on them and dragging. Also, it is possible to add new assignments by double clicking.

```{r, echo=FALSE, out.width='70%'}

if (html_output){
    gantt
} else {
    knitr::include_graphics('./../../img/calendar.png')
}
```

    Interactive gantt example

Here, an interactive graph is used to present the evolution of the status of a single aircraft over time. It is possible to filter, see the data in details and export as an image.

```{r, echo=FALSE, out.width='90%'}

if (html_output){
    graphs_rem
} else {
    knitr::include_graphics('./../../img/remaining.png')
}
```

    Interactive graph example

Here, an interactive table is used to present some data. It is possible to edit the contents, filter, re-order data, etc.

```{r, echo=FALSE, out.width='90%'}

if (html_output){
    table_rem
}
```

    Interactive table example

### Data integration

The application will include its own data storage through a database. All information used by the application will be stored here. In addition to this system, the application will permit several ways to import and export data from it.

#### Importing data

A new case can be created from the following options:

* The Excel file that is currently being used.
* An new yet-to-be-created Excel template file.

In addition to this, the historic data could be potential read from another system, if available.

#### Exporting data

The data in a case will be able to be exported in the following ways:

* In the *current Excel* format (or one that is similar).
* Into the new Excel template that is will be created.
* csv and/ or json files if needed.

This exporting capabilities will also permit the sharing of partial information to other actors in Excel or flat files.

### Architecture and deployment

The application will be deployed via a client-server architecture. This way, it would be easy to share the application given that it will only require having a modern web browser installed (Google Chrome).

The server should be installed somewhere inside the client's infrastructure.

The minimal hardware requirements for the server are:

* Ubuntu Server 64 bits.
* Cores >= 4
* RAM >= 8GB
* Hard drive >= 500 GB

The minimal software requirements for the server are:

* R + RStudio + Shiny
* python >= 3.6
* PostgreSQL

Additional software could be needed depending on the final specifications. For example, in case the model is way to big for open source solvers, a commercial solver might be needed.

Depending on the future integration, additional data flows will be included in the design. The only known data flow is between the existing Excel file and the application.

The following figure shows a summary of the architecture to be implemented and the data flows between components.

Users will interact via web browsers with the server. In order to load new data they can upload a new Excel file with the relevant data or edit an existing case via the application.

The server will be in charge of solving and generating new planning solutions and storing all relevant information in its database.

Users will be able to recover visually the results. Also, they will be able to export the results into csv, Excel and json formats depending on their needs.

```{r, echo=FALSE, out.width='70%'}

if (html_output){
    graph_architecture()

} else {
    knitr::include_graphics('../../img/architecture.png')
}
```

    Proposed architecture with data flows between components

This client-server architecture provides several advantages:

* Powerful graphics can be incorporated via open source javascript libraries.
* It is easier to maintain as a clear separation of concerns is done between GUI, logic and model.
* No installation of software besides a modern web browser is required.
* Multiple users can be easily incorporated.
* Integration with other systems can be done with minimal changes to the existing IT.
* Powerful resources can be allocated where needed (server) instead of the users PCs.

In addition to this, there is not third-party software to buy. All components are open source. The only possible exception could be the mathematical solver, depending on the size of the problem and the needs of speed.

### Users

Two possible roles are identified for the users of this application.

The first one will be called "planner". This user is in charge of creating (manually or automatically) different plannings.
The second one will be called "viewer". This user will be able to visualize the results of the planning created and stored inside the application. 
Depending on the needs of the client, several "viewer" profiles could be created in order to share only parts of the stored information with each profile.

## Team

<!-- TODO: names, CV and résumés-->

### Alain Haït

### Olga Battaïa

### Franco Peschiera

An industrial engineer, mathematician by profession and a programmer by heart.

Franco has been working in the Operations Research industry for almost 7 years. During which he has been the owner and Chief Technical Officer for a Spanish startup (www.baobabsoluciones.es) building software applications with mathematical optimization or simulation models inside for several big Spanish and international companies. Examples of clients have been: Repsol and CLH (oil logistics, marketing, human resources), General Motors, L.I. Lilly and Air Liquide (production planning), RATP, Alstom and ASM (logistics and transportation) among many others.

These projects involved the following incomplete list of tasks: the specification of the client"s requirements, the construction of a mathematical or simulation model, the programming of a graphical user interface, deployment and integration with the IT systems, documentation and training for users.

Franco started a PhD at ISAE-SUPAERO university in October 2017 on the topic of military aircraft maintenance planning.

## Planning

Here, the required tasks and their order are presented.
The main risk on the present planning is the incomplete information about:

* the specific problem to be solved.
* the available IT infrastructure to work with.
* availability of potential users to discuss example solutions and test and evaluate functionality.

### Phase 1: requirements

In this phase, several in-person and telephone meetings will be scheduled.

The requisites for this phase are: **None**

The deliverable of this phase will be:

* a document with an agreed description with the minimal requirements for the prototype and 
* a draft of possible additional requirements to be specified later on.

The duration of this phase is estimated to be: **3 weeks**

### Phase 2: first iteration

In this phase, a first version of the prototype will be constructed and deployed.

The requisites for this phase are:

* the list with minimal requirements.
* a deployment alternative to share the application with the users.

The deliverable of this phase will be:

* an prototype application that users can use and test.
* example solutions for users to analyze.

The duration of this phase is estimated to be: **5 weeks**

### Phase 3: second iteration

In this phase, a final version of the prototype will be constructed and deployed.

The requisites for this phase are:

* a list with changes to make to the first version based on the feedback from users.
* a list with additional functionality that has high priority.

The deliverable of this phase will be:

* an application that users can test.
* technical documentation.
* User's guide.

The duration of this phase is estimated to be: **5 weeks**

## Technologies and techniques

* **Mathematical Modeling** to solve the problem.
* **Python** for modeling, heuristics and data integration.
* **R** for statistics, possible forecasting and data visualization.
* **PostgreSQL** for storing data.
* **HTML5** for reports, visualization.

### Mathematical modeling

Mathematical modeling is a technique that is part of the branch of Operations Research in mathematics. It permits the construction of a set of equations to represent a specific well-defined situation and the use of advanced state of the art programs to search for and find the best solution that takes respects these equations. These programs (solvers) are available both in commercial and open source (free).

Advantages of using this technique compared to other alternatives are that it can prove the optimality of a solution, it can incorporate multiple objectives easily and it is possible to easily fix certain decisions that are already taken, making it possible to have interactive user-model use cases. With little modifications, it can be reused to take into account different situations.

Other functionality also includes having time limits and other more complex stop conditions.

### Python

python permits at the same time rapid prototyping as well as good code syntax. It offer a extremely wide offer of libraries and, in particular, it includes the best APIs to solver and non commercial solvers for Mixed Integer Programming problems.

It also excels at web developing with libraries such as flask and jinja templates.

### R

R is the state of the art in statistics and forecasting modelling. It includes several libraries for data manipulation, visualization and, as a programming language, good integration with other languages and software.

Finally, it offers specially good integration with html-based applications via the [Shiny](https://shiny.rstudio.com/gallery/) technology, which will be used in this project.

### PostgreSQL

PostgreSQL is one of the most well-known open-source database alternatives. It includes many functionalities that are not present in other database engines (namely MySQL) such as:

* NoSQL storing (for cases, for example).
* window-functions integrated with R.

### HTML5

HTML 5 is the most recent change to the HTML that incorporates graphical functionality. This standard has allowed a multitude of open sourced libraries that offer very powerful interactive graphics that are easily incorporated into webpages.

Examples of these libraries are: [leaflet](http://leafletjs.com/), [visjs](visjs.org), [d3js](https://d3js.org/) among many others.

## Disclaimers

1. The proposal presented above is intended for a complete application delivered as a part of a complete project to DGA.
1. The prototype to be delivered will not include the full functionality mentioned in this document but will show examples of the most important pieces of functionality following a meeting with the future users and decision makers.
1. Integration with other systems is out of the scope of this proposal since more information is needed regarding the existing systems to be able to estimate workload.
1. It is assumed that DGA will provide the hardware requirements together with the access rights and permissions to install all the components of the software. Alternatively, the software could be installed in the cloud upon agreement.
1. Even though mathematical programming provides the capacity to prove the optimality of a solution, the application does not guarantee finding it en every case. The quality of the solutions and the time to obtain them will depend on many factors; the size of the instance being the main one.

## References