### 2018-05-02 latex/notes/FMP_variants.tex
Add new bibliographie and notes on models. And CLAIO summary.

\documentclass[a4paper,11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}

%opening
\title{Variants on the Flight and Maintenance Planning problem}
\author{}

\setlength{\topmargin}{-2.1cm}
\setlength{\textwidth}{15.8cm}
\setlength{\textheight}{26.3cm}
\setlength{\oddsidemargin}{-0.1cm}
\setlength{\evensidemargin}{-0.1cm}

\def\II{{\mathcal{I}}}
\def\TT{{\mathcal{T}}}
\def\KK{{\mathcal{K}}}

\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{chngcntr}

\counterwithin*{equation}{section}
\counterwithin*{equation}{subsection}

\begin{document}

\maketitle

\section{Crew scheduling variant}

    As presented in \cite{Brunner2013}.

    \subsection{Hypothesis}

    \begin{itemize}
     \item fixed maintenance duration.
     \item fixed duration between maintenances.
     \item homogeneous fleet.
     \item objective: minimize the number of aircraft.
    \end{itemize}

    \subsection{Notations}

    \textbf{Sets}

    \begin{tabular}{ll}
    $t\in \mathcal{T} = [\![1, m]\!]$ & Planning horizon periods. \\
    $i \in \mathcal{P} = [\![1, n]\!]$ & Working pattern.\\
    \end{tabular}

    \vskip 0.3cm

    \subsection{Parameters}

    \begin{tabular}{ll}
        $D_{off}$ & Number of consecutive non-working days. \\
        $D_{on}$ & Number of consecutive working days. \\
        $\mathcal{A}_{it}$ & Pattern $i$ covers period $t$ (circulant matrix built from $D_{on}$ and $D_{off}$).\\
        $b_{t}$ & Need of resources at period $t$. \\
    \end{tabular}

    \vskip 0.3cm

    Example of A for $D_{on}=4$ and $D_{off}=2$

    $$
    \begin{bmatrix}
        1 & 0 & 0 & \dots  & A_{i1} \\
        1 & 1 & 0 & \dots  & A_{i2} \\
        1 & 1 & 1 & \dots  & A_{i3} \\
        1 & 1 & 1 & \dots  & A_{i4} \\
        0 & 1 & 1 & \dots  & A_{i5} \\
        0 & 0 & 1 & \dots  & A_{i6} \\
        1 & 0 & 0 & \dots  & A_{i6} \\
        \vdots & \vdots & \vdots & \ddots & \vdots \\
        A_{1t} & A_{2t} & A_{3t} & \dots  & A_{nm}
    \end{bmatrix}
    $$

    \subsection{Variables}

    \begin{itemize}
     \item $x_i$: number of patterns of type $i$ to be included in the solution.
    \end{itemize}

    \subsection{Model}

    \begin{align}
        & \text{Min}\; \sum_{i \in \mathcal{P}} x_i \label{numPatterns}
    \end{align}

    \begin{align}
        & \sum_{i \in \mathcal{P}} A_{it}x_{i} \geq d_j & t \in \mathcal{T} \label{demand} \\
        & x_{i} \geq 0,\; integer \\
    \end{align}

    (\ref{numPatterns}) minimizes the number of workers. (\ref{demand}) defines the need to have at least a given number of workers at any given time.

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Set covering variant}

    \subsection{Hypothesis}

    \begin{itemize}
     \item only 1 aircraft needed per month.
     \item objective: minimize the number of aircraft.
    \end{itemize}

    \subsection{Notations}

    \textbf{Sets}

    \begin{tabular}{ll}
    $t\in \mathcal{T} = [\![1, m]\!]$ & Planning horizon periods. \\
    $i \in \mathcal{P} = [\![1, n]\!]$ & Working pattern.\\
    \end{tabular}

    \vskip 0.3cm

    \subsection{Parameters}

    \begin{tabular}{ll}
        $\mathcal{A}_{it}$ & Pattern $i$ covers period $t$.\\
    \end{tabular}

    \vskip 0.3cm

    \subsection{Variables}

    \begin{itemize}
     \item $x_i$: is pattern $i$ in the solution or not.
    \end{itemize}

    \subsection{Model}

    \begin{align}
        & \text{Min}\; \sum_{i \in \mathcal{P}} x_i \label{numPatterns}
    \end{align}

    \begin{align}
        & \sum_{i \in \mathcal{P}} A_{it}x_{i} \geq 1 & t \in \mathcal{T} \label{demand} \\
        & x_{i} \geq 0,\; binary \\
    \end{align}

    (\ref{numPatterns}) minimizes the number of workers. (\ref{demand}) defines the need to have at least one worker working at any given time.

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Scheduling variant}

    \subsection{Hypothesis}

    \begin{itemize}
     \item durée de maintenance définie
     \item fenêtres de temps pour commencer les maintenances
     \item objectif: lisser les maintenances dans les fenêtres de temps prédéfinies
    \end{itemize}

    \subsection{Notations}

    \noindent{
    \begin{tabular}{ll}
        $t\in \TT = [\![1, T]\!]$ & Discrétisation temporelle de l'horizon de planigfication. \\
        $i \in \II = [\![1, I]\!]$ & Ensemble des avions.\\
        $k \in \KK_i = [\![1, K_i]\!]$ & Ensemble des maintenances à planifier successivement.\\ 
    \end{tabular}
    }

    \vskip 0.3cm

    \textbf{Paramètres}

    \begin{tabular}{ll}
        $\mathbf{Da}_{i,k}$ & Durée nominale de la maintenance $k$ de l'avion $i$.\\
        $\mathbf{l}_{i}$ & Durée mininale du cycle de vol de l'avion $i$.\\
        $\mathbf{{To}}_{i,k}$ & Pas de temps le plus tôt pour commencer la maintenance $k$ de l'avion $i$.\\
        $\mathbf{{Ta}}_{i,k}$ & Pas de temps le plus tard pour terminer la maintenance $k$ de l'avion $i$.\\
    \end{tabular}

    % \caption{Definition and notations the challenge ROADEF}\label{contraintesROADEF}
    % \end{table}

    \subsection{Définition des variables}

    \begin{itemize}
     \item $d_{i,k,t}\in \{0,1\}$; variables binaires pour indiquer les dates d'arrêts, "step varaibles" 
    $d_{i,k,t}=1$ si l'arrêt du cycle $k$ du réacteur nucléaire $i$ a été commencé avant la semaine $t$.
    $d_{i,k,t}-d_{i,k,t- {\mathbf{Da}_{i,k}}} \in \{0,1\}$ vaut $1$ uniquement pendant a maintenance $(i,k)$.

     \item $m,M\geqslant 0$: variables continues pour linéariser la fonction objectif
    \end{itemize}

    \subsection{Formulation PLNE}

    \begin{eqnarray}
    \displaystyle \min_{d,m,M} & M-m
     & \\
    \forall i , k, t, & d_{i,k,t-1}\leqslant d_{i,k,t} & \label{PANprecedence}\\
    \forall i , k, & d_{i,k,\mathbf{To}_{i,k}-1}\leqslant 0 & \label{PANtw0} \\
    \forall i , k, & d_{i,k,\mathbf{Ta}_{i,k}}\geqslant 1 & \label{PANtw1}\\
    \forall i , k, t, & d_{i,k,t}\leqslant d_{i,k-1,t-l_i-\mathbf{Da}_{i,k}} & \label{PANprecedence2}\\
    \forall t, & \displaystyle \sum_{i,k} d_{i,k,t}-d_{i,k,t- {\mathbf{Da}_{i,k}}} \leqslant M & \label{maxMaint}\\
    \forall t, & \displaystyle \sum_{i,k} d_{i,k,t}-d_{i,k,t- {\mathbf{Da}_{i,k}}} \geqslant m & \label{minMaint}\\
    & d \in \{0,1\}^N, m,M \geqslant 0
    \end{eqnarray}

    (\ref{PANprecedence}): contrainte induite par la définition ``step variables".

    (\ref{PANtw0}), (\ref{PANtw1}): contraintes de fenêtres de temps (écrit explicitement, ou sinon définir uniquement les variables $d_{i,k,t}$
    dans les fenêtres de temps)

    (\ref{PANprecedence2}): durée minimale d'un cycle de vol $l\geqslant 0$, contrainte de succession des cycles.

    (\ref{maxMaint}), (\ref{minMaint}): contraites pour linéariser les $\min$ et $\max$ de la focntion objectif de lissage

    \subsection{Questions ouvertes}

    \begin{itemize}
    %\item Avec $l=0$ , y a t'il un problème de référence proche ? Ordonnancement, lissage de charge ?
    \item Avec $l=0$ , y a t'il un problème de référence proche ? Ordonnancement, lissage de charge ?
    \item Est-ce un problème NP-complet? Peut on le prouver ?
     \item Cas particulier $\mathbf{Ta}_{i,k}-\mathbf{To}_{i,k}=1$? (pour lissage local)
     \item Autres fonctions objectifs?
     \item version robuste avec aléas de prolongation de durée de maintenance?
    \end{itemize}

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Crew scheduling: smooth maintenances}

    \subsection{Hypothesis}

    \begin{itemize}
     \item fixed maintenance duration.
     \item range to start each maintenance.
     \item min distance between maintenances.
     \item objective: minimize the max number of maintenances.
    \end{itemize}

    \subsection{Notations}

    \textbf{Sets}

    \begin{tabular}{ll}
    $t\in \mathcal{T} = [\![1, m]\!]$ & Planning horizon periods. \\
    $i \in \mathcal{P} = [\![1, n]\!]$ & Working pattern.\\
    \end{tabular}

    \vskip 0.3cm

    \subsection{Parameters}

    \begin{tabular}{ll}
        $\mathcal{A}_{it}$ & Pattern $i$ has a maintenance in $t$.\\
    \end{tabular}

    \vskip 0.3cm

    Here, A needs to be created taking into account:

    \begin{enumerate}
        \item The range of start of maintenances.
        \item The fixed duration of maintenances.
        \item The minimal distance between maintenances.
    \end{enumerate}

    \subsection{Variables}

    \begin{itemize}
        \item $x_i$: number of patterns of type $i$ to be included in the solution.
        \item $M$: maximum number of maintenance among all periods.
    \end{itemize}

    \subsection{Model}

    \begin{align}
        & \text{Min}\; M
    \end{align}

    \begin{align}
        & \sum_{i \in \mathcal{P}} A_{it}x_{i} \leq M & t \in \mathcal{T} \label{numMaint} \\
        & x_{i} \geq 0,\; integer \\
    \end{align}

        (\ref{numMaint}) defines the $M$ variable as the maximum number of maintenances at any period of time.

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Scheduling variant 2}

    As modelled in \cite{Rieck2012} and \cite{Neumann2000}.

    \subsection{Hypothesis}

    \begin{itemize}
     \item fixed maintenance duration.
     \item min and max durations between maintenances that belong to the same aircraft.
     \item each maintenance uses an amount of resource. In this case 1.
     \item objective: minimize the max number of maintenances.
     \item There are two artificial activities representing the start and end of the horizon.
    \end{itemize}

    \subsection{Notations}

    \textbf{Sets}

    \noindent{
    \begin{tabular}{ll}
        $i \in \mathcal{V} = [\![1, n]\!]$ & Set of maintenance operations.\\
        $(i, j) \in \mathcal{E} = [\![1, v]\!]$ & Set of consecutive (i, j) maintenance operations, i.e. consecutive maintenances belonging to the same aircraft.\\
        $k \in \mathcal{R}$ & Types of resources that maintenance can take. We only have one type.\\
        $i = 0$ & Artificial starting node.\\
        $i = n+1$ & Artificial ending node.\\
    \end{tabular}
    }

    \vskip 0.3cm

    The set $\mathcal{A}$ we will construct it by deciding in advance the maintenances $i$ that share the same aircraft.

    \subsection{Parameters}

    \begin{tabular}{ll}
        $p_i$ & Duration of maintenance $i$.\\
        $\mathbf{d}$ & Last period on the horizon.\\
        $r_{ik}$ & Resources of type $k$ taken by maintenance $i$. In this case it's always 1 per maintenance.\\
        $d^{max}_{ij}$ & Maximum duration between two maintenances $i$ and $j$.\\
        $d^{min}_{ij}$ & Minimum duration between two maintenances $i$ and $j$.\\
        ${\delta}_{ij}$ & Precedence duration arcs between two maintenances $i$ and $j$ based on $d^{max}$ and $d^{min}$.\\
        $R_k$ & total maintenance capacity at each period for resource $k$.\\
        $c_k$ & cost per unit of investment on resource $k$.
        % $\mathbf{{To}}_{i,k}$ & Minimum starting period for maintenance $k$ of resource $i$.\\
        % $\mathbf{{Ta}}_{i,k}$ & Maximum starting period for maintenance $k$ of resource $i$.\\
    \end{tabular}

    \subsection{Variables}

    \begin{tabular}{ll}
        $S_i$ & the start of maintenance operation $i$.\\
        $\mathcal{A}(S^{\mathcal{C}}, t)$ & set of maintenance operations active at moment $t$.\\
    \end{tabular}

    \subsection{Model}

    The following formulation is taken from \cite{Neumann2000}.

    \begin{align}
        \min f(\mathbf{R}(S)) = \sum_{k \in \mathcal{R}} c_k \max_{t \in \mathcal{T}} r_k(S^{\mathcal{C}}, t)
    \end{align}

    \begin{align}
        & S_j - S_i \geq \delta_{ij} & (i, j) \in \mathcal{E}  \label{precedence}\\
        & S_0 = 0 &  \label{firstjob}\\
        & S_{n+1} \leq \mathbf{d}   \label{lastjob}\\
        & S_{i} \in \mathbb{Z}_{\geq 0} & i \in \mathcal{V}  \label{integervars}\\
        & r_k(S^{\mathcal{C}}, t) = \sum_{i \in \mathcal{A}(S^{\mathcal{C}}, t)} r_{ik}  \label{consumptiondef}\\
        & r_k(S^{\mathcal{C}}, t) \leq R_k \label{consumptionmax}\\
    \end{align}

    (\ref{precedence}) constraints define the relative minimum and maximum starting times of tasks.
    (\ref{firstjob}) and (\ref{lastjob}) constraints define the fictitious start and end jobs.
    (\ref{integervars}) define the range for the starting time variables: integer positive.
    (\ref{consumptiondef}) defines the used quantity for each type of resource depending on the starting times and durations and (\ref{consumptionmax}) defines the maximum amount of resources of each type that can be used at any given time.

% As it has been shown by Bartusch et al. (1988),
% the feasible region S (without the integrality
% constraints S i 2 Z, i 2 V ) of R   is generally dis-
% connected and represents the union of (convex)
% polytopes whose number grows exponentially in n.
% The decision problem whether or not S 6 ;

% \bibliographystyle{dcu}
\bibliographystyle{alpha}
\bibliography{./../biblio/Scheduling,./../biblio/NRP}

\end{document}