### 2018-06-01 R/diagrams/heuristic/heuristic_2.Rmd

rut = remaining usage time = remaining flight hours.
ret = remaining elapsed time = remaining periods.

------

**The base logic** (probably similar to what's in the students' PIE code):

Sort *periods* from lowest to greatest.
For each period:
    Get *tasks* that need at least one aircraft in that period.
    Sort *tasks* according to the number of candidates they have.
    For each task in *tasks*:
        Get available *candidates* that can do the task.
        Sort *candidates* according to 'tie methodology' (largest $rut / ret$ first).
        For each candidate:
            Try to assign candidate to task in period.
                If possible: update the candidate's rut and task remaining needs.
                If not possible: give a maintenance to the candidate.
            if no more remaining needs => go to the next task.

**The additional logic** (that should be executed at the end of the each period loop):

We simulate the next X (10?) periods only by using ret to generate maintenances (since we're using the 'tie methodology', this could be "good enough").

if this simulation:

* shows periods with too many maintenances: 
    * we try to advance some of those maintenances to the present. 
    * (How many?) The number of maintenances to advance may depend on how many maintenances we can actually do and how many months we can move the maintenances.
* shows periods with non-availabilities for missions: 
    * we try to advance some maintenances in that period from an aircraft with the same capabilities to the present. 
    * (How many?) This can be the same logic than the one before.

Other comments on these 'moves':

* We should first try to advance the aircraft that have the least ret.
* We need to check that the changed periods continue to be feasible.
* The changed periods are the ones at each extreme of the maintenance and the number will be two times the size of the move.