### 2018-09-19 R/presentations/CLAIO_presentation/presentation.Rmd
second version of CLAIO ppt

---
title: "Maintenance planning on French military aircraft operations"
author: |
    | Franco Peschiera, Alain Haït, Olga Battaïa, Nicolas Dupin
institute: "ISAE SUPAERO"
date: "September 24, 2018"
output: 
  slidy_presentation:
    css: ./../../../css/baobab_slidy.css
    incremental: yes
    footer: "ISAE-SUPAERO, Université de Toulouse, France"
    # slide_level: 2
    # toc: true
  beamer_presentation:
    theme: "Berlin"
    toc: true
    slide_level: 2
    incremental: yes
    fig_height: 4
    # citation_package: natbib
    # footer: "ISAE-SUPAERO, Université de Toulouse, France"
    includes:
        in_header: header.tex
bibliography: MFMP_NRP.bib
# nocite: '@*'
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
path_to_R_dir <- "./../../"
setwd(path_to_R_dir)
source('functions/params.R')
source('functions/import_results.R')
source('functions/scenario_analysis.R')
source('diagrams/fixed_interval.R')
all_exp_dir <- PATHS[['results']] %>% paste0('simulated_data/')
exp_directory <- all_exp_dir %>% paste0('2_task/201809180941/')
gantt <- print_solution(exp_directory, max_resources = 15, width='100%')
show_gantt <- knitr::opts_knit$get("rmarkdown.pandoc.to")=="slidy"
task_gantt <- tasks_gantt()
task_employees <- tasks_employees()
scen_means <- scenario_mean(all_exp_dir)
infeasible <- scenario_get_gaps(all_exp_dir) %>% lapply(function(scn){
    a <- length(scn)
    scn %>% is.na %>% sum %>% divide_by(a/100)
} )

```

# Problem

## Problem (informally)

Assign both missions and maintenance operations to a fleet of aircraft in order to maximize availability and minimize costs. Missions have fixed start and end times and have particular needs in terms of aircraft and time.

## Problem (example)

```{r, echo=FALSE, out.width='70%'}

if (show_gantt){
    gantt
} else {
    knitr::include_graphics('../img/calendar.png' %>% paste0(path_to_R_dir, .))
}

```

## Problem

* FMP: Flight and Maintenance Planning problem.
* A series of $j \in \mathcal{J}$ tasks are planned along a horizon divided into $t \in \mathcal{T}$ periods. Since all tasks are already scheduled, we know time periods $T_j \subset \mathcal{T}$ in which they will be realized. 
* Each task requires a certain number $R_{j}$ of resources $i \in \mathcal{I}$ which it employs for a time duration defined by $H_j$ in each period and needs to be assigned for a minimum of $MT_j$ consecutive periods if assigned.
* Resources require recurrent preventive maintenance operations since the realization of tasks diminish their remaining usage time.
* A maintenance operation takes exactly $m$ periods and cannot be interrupted. It restores the resource's remaining usage time to exactly $H^M$ units.

# State of the art

## State of the art (1)

* In @Cho2011, US Army aircraft were assigned daily operations over a year to minimize the number of maintenances.
* In @Kozanidis2008, Greek aircraft had monthly assignments of maintenances and flight hours in order to maximize the availability and final state of squadrons.
* In @Verhoeff2015, monthly assignments were done and several objectives were taken into account: availability, serviceability and final state.
* In @Seif2018, a generalization of @Kozanidis2008 was done in order to deal with different types of maintenances and an heterogeneus fleet.

## State of the art (2)

* Asignments to missions are done instead of assignments of flight hours.
* Missions need to be assigned for a minimal duration.
* This model also features several improvements in the MIP modelling that permit greater instance sizes.

# Complexity analysis

## Complexity analysis

<div class="incremental">

Reduction to the 'Fixed interval scheduling' in @Smet2015 **proves NP-completeness**.

* $P$ tasks with duration $u_p$ and start time $s_p$.
* $E$ employees with a set of tasks $P_e \in P$ that each can perform.
* Minimize the number of required employees to do all tasks.

\pause

```{r, echo=FALSE, out.width='45%'}

if (show_gantt){
    htmltools::div(class = "row",
        htmltools::div(task_employees, class = "col-md-6")
        ,htmltools::div(task_gantt, class = "col-md-6")
        )
} else {
    c('../img/tasks-days.png', '../img/task-employees.png') %>% 
        paste0(path_to_R_dir, .) %>%
        knitr::include_graphics()
    # knitr::include_graphics('../img/task-employees.png' %>% paste0(path_to_R_dir, .))
}

```

</div>

# Model

## Model: main constraints

<div class="incremental">

\begin{align}
& \text{Min}\; \sum_{t \in \mathcal{T}, i \in \mathcal{I}} m_{it} \times H^{M} - \sum_{i \in \mathcal{I}} rut_{i|\mathcal{T}|}
\end{align}

The objective is to minimize the total number of maintenances and, at the same time, maximize the end status of the resources.

\pause

\begin{align}
    & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \leq C^{max}
      & t \in \mathcal{T} \label{eq:capacity1}\\
    & \sum_{i \in \mathcal{I}_j} a_{jti} \geq R_j
            & j \in \mathcal{J}, t \in \mathcal{T}_j  \label{eq:taskres}\\
    & \sum_{t' \in \mathcal{T}^{s}_t} m_{it'} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} \leq 1 
            & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}
\end{align}

Mission's needs and incompatibility of several tasks or maintenance in the same period.

</div>

## Model: maintenances

\begin{align}
    & m_{it'} + m_{it} \leq 1
      & t \in \mathcal{T}, t' \in \mathcal{T}^{m}_t, i \in \mathcal{I}\label{eq:ret_min}\\ 
    & \sum_{t' \in \mathcal{T}^{M}_t} m_{it'} \geq  m_{it}
      & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_max}\\
    & m_{it} = 0
      & t \in \mathcal{T}^{m_{ini}}_i, i \in \mathcal{I}\label{eq:ret_min_init} \\
    & \sum_{t \in \mathcal{T}^{M_{ini}}_i} m_{it} \geq  1 
      & i \in \mathcal{I}\label{eq:ret_max_init}
\end{align}

Maintenances cannot executed too soon or too late with respect to the previous one.

<!-- </div> -->

<!-- ## Model: time consumption -->

<!-- Time usage of an aircraft depending on the assignment to a task and remaining usage time per aircraft and period. -->

<!-- \begin{align} -->
<!--  & u_{it} \geq \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j  -->
<!--     & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:usage1}\\ -->
<!--  & u_{it} \geq u^{min} (1 - \sum_{t' \in \mathcal{T}^{s}_t} m_{it'}) -->
<!--     & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:usage2}\\ -->
<!--  & u_{it} \in [0, \max_j{\{H_j\}}] -->
<!--     & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:usage_lower}\\ -->
<!--  & rut_{it} \leq rut_{i(t-1)} + H^{M} m_{it} - u_{it} -->
<!--     & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rut_upper}\\ -->
<!--  & rut_{i0} = Rut^{Init}_i -->
<!--        & i \in \mathcal{I} \label{eq:rut_initial}\\ -->
<!--  & rut_{it} \geq H^{M} m_{it'} -->
<!--         & t \in \mathcal{T}, t' \in \mathcal{T}^{s}_t, i \in \mathcal{I}\label{eq:rut_lower}\\ -->
<!--  & rut_{it} \in [0,H^{M}] -->
<!--         & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:mu} -->
<!-- \end{align} -->

## Model: other constraints

* Minimum cluster availability.
* Minimum duration for task assignment.
* Hour consumption of resources depending on tasks and maintenances.

<!-- ## Model: availability -->

<!-- \begin{align} -->
<!--    & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}_k} m_{it'} + N^{clust}_{kt} \leq A^{clust}_{kt} -->
<!--     &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:avalaibility-cluster} \\ -->
<!--    & \sum_{i \in \mathcal{I}_k} rut_{it} \geq H^{clust}_{kt} -->
<!--     &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:avalaibility-hours-cluster} -->
<!-- \end{align} -->

<!-- Each cluster of aircraft needs a minimum level of availability in number and usage hours. -->

# Results and perspectives

## Experiments

<div class="incremental">

* 60 - 200 resources.
* 90 periods.
* 3 - 9 tasks:
    * 12 - 36 periods long
    * 2 - 5 resources needed

\pause

* Small instances have **30.000 variables** and **20.000 constraints**.
* Medium instances have **80.000 variables** and **50.000 constraints**.
* Runs of 30 minutes have been done on 30 randomly generated cases per scenario.
* CPLEX was used in an i7, quadcore Ubuntu 18.04 workstation.

</div>

## Statistics

* `r infeasible[1]`% of instances were proved infeasible inmediately.
* From feasible small instances, a mean gap of `r scen_means[1]`% was found after 30 min.
* From feasible medium instances, a mean gap of `r scen_means[2]`% was found after 30 min.

## Perspectives

* More extreme (bigger) scenarios: more tasks and more resources.
* A deeper analysis on the detailed solving process.
* Benchmarking of solvers (gurobi, CBC, choco and CPO).
* Heuristics for large scale instances.
* The inclusion of new possibilities, such as the planned storage of resources.
* A stochastic sensibility analysis, including taking account of stochasticity in the input data.

# References

## References / Questions?

<!-- \bibliographystyle{dcu} -->
<!-- \bibliography{MFMP_NRP.bib} -->