### 2018-11-30 R/presentations/CLAIO_presentation/presentation.Rmd

---
title: "Maintenance planning on French military aircraft operations"
author: |
    | Franco Peschiera, Alain Haït, Olga Battaïa, Nicolas Dupin
institute: "ISAE-SUPAERO, Université de Toulouse, France"
date: "September 24, 2018"
output: 
  beamer_presentation:
    # keep_tex: true
    theme: "Berlin"
    toc: true
    slide_level: 2
    incremental: yes
    fig_height: 4
    # citation_package: natbib
    # footer: "ISAE-SUPAERO, Université de Toulouse, France"
    includes:
        in_header: header.tex
  slidy_presentation:
    css: ./../../../css/baobab_slidy.css
    incremental: yes
    footer: "ISAE-SUPAERO, Université de Toulouse, France"
    # slide_level: 2
    # toc: true
bibliography: MFMP_NRP.bib
# nocite: '@*'
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
path_to_R_dir <- '.'
path_to_R_dir <- "./../../"
setwd(path_to_R_dir)
source('functions/params.R')
source('functions/import_results.R')
source('functions/scenario_analysis.R')
source('diagrams/fixed_interval.R')
all_exp_dir <- PATHS[['results']] %>% paste0('simulated_data/')
exp_directory <- all_exp_dir %>% paste0('task_periods_solv_2_50_GUROBI/201810121218/')
gantt <- print_solution(exp_directory, max_resources = 15, width='100%')
show_gantt <- knitr::opts_knit$get("rmarkdown.pandoc.to")=="slidy"
task_gantt <- tasks_gantt()
task_employees <- tasks_employees()
scen_means <- scenario_mean(all_exp_dir)
infeasible <- scenario_get_gaps(all_exp_dir) %>% lapply(function(scn){
    a <- length(scn)
    scn %>% is.na %>% sum %>% divide_by(a/100)
} )

```

# Problem

## Problem (informally)

**Assign** both **missions and maintenance** operations to a fleet of **aircraft** in order to maximize **availability** and minimize **costs**. **Missions** have **fixed** start and end times and have particular **needs** in terms of aircraft and time.

## Problem (example)

```{r, echo=FALSE, out.width='70%'}

if (show_gantt){
    gantt
} else {
    knitr::include_graphics('../img/calendar.png' %>% paste0(path_to_R_dir, .))
}

```

## Problem

* FMP: Flight and Maintenance Planning problem.
* A series of $j \in \mathcal{J}$ tasks are planned along a horizon divided into $t \in \mathcal{T}$ periods.
* Each task requires $R_{j}$ of resources for $H_j$ hours and needs to be assigned for a minimum of $MT_j$ consecutive periods.
* Resources $i \in \mathcal{I}$ require recurrent preventive maintenance operations.
* A maintenance operation takes exactly $m$ periods and restores the resource's remaining usage time to exactly $H^M$ units.

# State of the art

## State of the art (1)

* In @Cho2011, US Air Force aircraft were assigned daily operations over a year to minimize the number of maintenances.
* In @Kozanidis2008, Greek aircraft had monthly assignments of maintenances and flight hours in order to maximize the availability and final state of squadrons.
* In @Verhoeff2015, monthly assignments were done and several objectives were taken into account: availability, serviceability and final state.
* In @Seif2018, a generalization of @Kozanidis2008 was done in order to deal with different types of maintenances and an heterogeneus fleet.

## State of the art (2)

* Asignments to missions are done instead of assignments of flight hours.
* Missions need to be assigned for a minimal duration.
* Because of these differences, a new model is proposed to deal with this new problem.

# Complexity analysis

## Complexity analysis

<div class="incremental">

Reduction to the 'Fixed interval scheduling' in @Smet2015 **proves NP-completeness**.

* $P$ tasks with duration $u_p$ and start time $s_p$.
* $E$ employees with a set of tasks $P_e \in P$ that each can perform.
* Minimize the number of required employees to do all tasks.

\pause

```{r, echo=FALSE, out.width='45%'}

if (show_gantt){
    htmltools::div(class = "row",
        htmltools::div(task_employees, class = "col-md-6")
        ,htmltools::div(task_gantt, class = "col-md-6")
        )
} else {
    c('../img/tasks-days.png', '../img/task-employees.png') %>% 
        paste0(path_to_R_dir, .) %>%
        knitr::include_graphics()
    # knitr::include_graphics('../img/task-employees.png' %>% paste0(path_to_R_dir, .))
}

```

</div>

# Model

## Model: variables

The following variables are used in the model.

\begin{tabular}{lll}
$m_{it}$    &: maintenance start &  $|\mathcal{I}| \times |\mathcal{T}|$ \\
$a_{jti}$   &: task assignment &  $|\mathcal{T}| \times |\mathcal{J}_t| \times |\mathcal{I}_j|$ \\
$a^s_{jti}$ &: task assignment start &  $|\mathcal{T}| \times |\mathcal{J}_t| \times |\mathcal{I}_j|$ \\
$u_{it}$    &: monthly usage time &  $|\mathcal{I}| \times |\mathcal{T}|$ \\
$rut_{it}$  &: remaining usage time &  $|\mathcal{I}| \times |\mathcal{T}|$ \\
\end{tabular}

## Model: main constraints

<div class="incremental">

\begin{align}
& \text{Min}\; \sum_{t \in \mathcal{T}, i \in \mathcal{I}} m_{it} \times H^{M} - \sum_{i \in \mathcal{I}} rut_{i|\mathcal{T}|}
\end{align}

The objective is to minimize the total number of maintenances and, at the same time, maximize the end status of the resources.

\pause

\begin{align}
    & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \leq C^{max}
      & t \in \mathcal{T} \label{eq:capacity1}\\
    & \sum_{i \in \mathcal{I}_j} a_{jti} \geq R_j
            & j \in \mathcal{J}, t \in \mathcal{T}_j  \label{eq:taskres}\\
    & \sum_{t' \in \mathcal{T}^{s}_t} m_{it'} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} \leq 1 
            & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}
\end{align}

Mission's needs and incompatibility of several tasks or maintenance in the same period.

</div>

## Model: maintenances

\begin{align}
    & m_{it'} + m_{it} \leq 1
      & t \in \mathcal{T}, t' \in \mathcal{T}^{m}_t, i \in \mathcal{I}\label{eq:ret_min}\\ 
    & \sum_{t' \in \mathcal{T}^{M}_t} m_{it'} \geq  m_{it}
      & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_max}
\end{align}

<!-- Maintenances cannot executed too soon or too late with respect to the previous one. -->

```{r, echo=FALSE, out.width='60%', fig.align="center"}
'../img/rut_ret.png' %>% 
    paste0(path_to_R_dir, .) %>%
    knitr::include_graphics()
```

<!-- </div> -->

<!-- ## Model: time consumption -->

<!-- Time usage of an aircraft depending on the assignment to a task and remaining usage time per aircraft and period. -->

<!-- \begin{align} -->
<!--  & u_{it} \geq \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j  -->
<!--     & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:usage1}\\ -->
<!--  & u_{it} \geq u^{min} (1 - \sum_{t' \in \mathcal{T}^{s}_t} m_{it'}) -->
<!--     & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:usage2}\\ -->
<!--  & u_{it} \in [0, \max_j{\{H_j\}}] -->
<!--     & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:usage_lower}\\ -->
<!--  & rut_{it} \leq rut_{i(t-1)} + H^{M} m_{it} - u_{it} -->
<!--     & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rut_upper}\\ -->
<!--  & rut_{i0} = Rut^{Init}_i -->
<!--        & i \in \mathcal{I} \label{eq:rut_initial}\\ -->
<!--  & rut_{it} \geq H^{M} m_{it'} -->
<!--         & t \in \mathcal{T}, t' \in \mathcal{T}^{s}_t, i \in \mathcal{I}\label{eq:rut_lower}\\ -->
<!--  & rut_{it} \in [0,H^{M}] -->
<!--         & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:mu} -->
<!-- \end{align} -->

## Model: other constraints

* Minimum cluster availability.
* Minimum duration for task assignment.
* Hour consumption of resources depending on tasks and maintenances.

<!-- ## Model: availability -->

<!-- \begin{align} -->
<!--    & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}_k} m_{it'} + N^{clust}_{kt} \leq A^{clust}_{kt} -->
<!--     &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:avalaibility-cluster} \\ -->
<!--    & \sum_{i \in \mathcal{I}_k} rut_{it} \geq H^{clust}_{kt} -->
<!--     &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:avalaibility-hours-cluster} -->
<!-- \end{align} -->

<!-- Each cluster of aircraft needs a minimum level of availability in number and usage hours. -->

# Results and perspectives

## Example of a solution

```{r, echo=FALSE, out.width='100%', fig.align="center"}
'../img/example_solution1.png' %>% 
    paste0(path_to_R_dir, .) %>%
    knitr::include_graphics()
```

## Results on previous instances

All instances are solved to 0.02% of optimality in less than one hour. Previously this was not the case.

| id | $\|\mathcal{J}\|$ |$\|\mathcal{T}\|$ |  assign |  gap (\%)   |  time (s) |
|-----|------------------|------------------|---------|-------------|-----------|
| I_0 |                9 |               11 |     310 |      0.01 	|     29.79 |
| I_1 |                9 |               21 |     650 |      0.02 	|    173.11 |
| I_2 |                9 |               31 |     990 |      0.02 	|    409.66 |
| I_3 |                9 |               41 |    1249 |      0.03 	|    836.08 |
| I_4 |               10 |               11 |     530 |      0.02 	|     37.66 |
| I_5 |               10 |               21 |    1070 |      0.01 	|    208.25 |
| I_6 |               10 |               31 |    1610 |      0.01 	|    331.90 |
| I_7 |               10 |               41 |    2069 |      0.02 	|   1731.50 |
| I_8 |               11 |               11 |    1080 |      0.01 	|     76.18 |
| I_9 |               11 |               21 |    2120 |      0.02 	|   3332.95 |

## Experiments

Random instances were generated to test the performance following the known instances guidelines.

<div class="incremental">

* 60 - 200 resources.
* 90 periods.
* 3 - 9 tasks:
    * 12 - 36 periods long
    * 2 - 5 resources needed

\pause

* Instances have **30.000 - 80.000 variables** and **20.000 - 50.000 constraints**.
* Runs of 30 minutes have been done on 30 randomly generated cases per scenario.
* CPLEX was used in an i7, quadcore Ubuntu 18.04 workstation.

</div>

## Statistics

* From feasible small instances, a mean gap of `r scen_means[1]`% was found after 30 min.
* From feasible large instances, a mean gap of `r scen_means[2]`% was found after 30 min.

## Perspectives

* More extreme (bigger) scenarios: more tasks and more resources.
* A deeper analysis on the detailed solving process.
* Benchmarking of solvers (gurobi, CBC, choco and CPO).
* Heuristics for large scale instances.
* The inclusion of new possibilities, such as the planned storage of resources.
* A stochastic sensibility analysis, including taking account of stochasticity in the input data.

# References

## References / Questions?

<!-- \bibliographystyle{dcu} -->
<!-- \bibliography{MFMP_NRP.bib} -->