### 2018-11-30 latex/CLAIO2018/article.tex

%========================================================
% File:         claio_paper.tex, template for CLAIO 2015 conference paper
% Based on:     CLAIO 2014 paper template
%========================================================
\documentclass[letterpaper,11pt]{article}

% added by Roger
% \usepackage{psfig,epsfig}
\usepackage{layout}
% added by J. Vera
\usepackage[pdftex]{graphics}  % In case you want to use PDFTeX
\usepackage{graphicx}
\renewcommand{\baselinestretch}{1.0}
\usepackage{amsmath}
\usepackage[latin1]{inputenc}
\usepackage[frenchb,english]{babel}
%
%       Text dimensions.
%       PLEASE, DO NOT CHANGE THE TEXT DIMENSIONS.
%
\setlength{\textwidth}{6.5in}
\setlength{\textheight}{8.5in}
\setlength{\oddsidemargin}{0.0in}
\setlength{\evensidemargin}{0.0in}
\setlength{\topmargin}{0.0in}
\setlength{\headsep}{0.1in}
\setlength{\headheight}{0.4in}

\pagestyle{empty}

%
%  Title definition
%
\title{\bf A novel MIP formulation for the optimization problem of maintenance planning of military aircraft}

\author{%
ISAE-SUPAERO, Université de Toulouse, France\\%
Franco Peschiera\\%
\textit{franco.peschiera@isae-supaero}\\[3mm]
Nicolas Dupin\\%
\textit{nicolas.dupin@isae-supaero}\\[3mm]
Olga Battaïa\\%
\textit{olga.battaia@isae-supaero}\\[3mm]
Alain Haït\\%
\textit{alain.hait@isae-supaero}
}

\date{ }

%
%  Document starts here
%
\begin{document}

\maketitle

\thispagestyle{empty}

\begin{abstract}

{\em A Flight and Maintenance Planning problem is considered for military aircraft. A novel Mixed-Integer Linear Programming model is developed for scheduling maintenance operations along with the assignment of regular, pre-scheduled, missions. This model takes into account the specific requirements of a military environment and its operational goals. On the contrary to previous studies, the developed model is applicable for the management of a heterogenic fleet. Different combinations of objective functions are discussed for multi-objective optimization and the impact of their use is analyzed on a dataset provided by French Army Forces. The complexity of the obtained problem formulations is studied; new procedures are developed in order to decrease the computational time required to construct efficient Pareto-fronts of alternative solutions for real-life sized instances. The results obtained in computational tests show the efficiency of the developed optimization approach. 
}
\vspace{2em}

\noindent\textit{Keywords:}
optimization;
Flight and Maintenance Planning;
aircraft maintenance;
Mixed Integer Programming.

\end{abstract}
\end{document}