### 2019-08-19 R/scripts/resultsNPS.Rmd
added markdown for results reporting.

---
title: "ResultsNPS"
author: "Franco Peschiera"
date: "August 8, 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(reticulate)
use_python('C:/Python36/python.exe')
use_virtualenv('C:/Users/pchtsp/Documents/projects/optima/python/venv/', required = TRUE)
current_dir = getwd()
print(current_dir)
# Sys.getenv('PYTHONHOME')
# # Sys.setenv(PYTHONPATH='C:/Python36/DLLs')
# py_available(TRUE)
py_discover_config()

```

```{python}
import pandas as pd
import numpy as np

import sys, os
path = os.path.join(r.current_dir, './../../python/')
print(path)
sys.path.insert(1, path)

import package.batch as ba
# import package.params as params

# from rpy2.robjects import pandas2ri
# import rpy2.robjects.lib.ggplot2 as ggplot2
# import rpy2.robjects as ro

# import stochastic.params as sto_params

```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.