### 2019-09-10 R/presentations/DISC_201909/presentation.Rmd
added presentation to DISC, many changes in R reports and graphs, some code corrections on batch instances in python

---
title: "Predicting maintenance patterns"
author: "Franco Peschiera"
date: "September, 2019"
output:
  xaringan::moon_reader:
    self_contained: false
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
  slidy_presentation:
    css: ./../../../css/baobab_slidy.css
    incremental: yes
    footer: "ISAE-SUPAERO, Université de Toulouse, France"
    slide_level: 3
    toc: true
editor_options: 
  chunk_output_type: console
---

### Structure of talk

1. Decision problem.
2. Prediction problem.
2. Current results.
3. About my ambition.

---

## Decision problem

---

### Problem (informally)

Assign both **missions** and **maintenance operations** to a fleet of aircraft in order to maximize availability and minimize costs. **Missions** have fixed start and end times and have particular needs in terms of aircraft and flight hours. **Maintenances operations** have specific rules that govern their frequency and capacity.

--

.center[
```{r, echo=FALSE, out.width='100%'}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
knitr::include_graphics('./../../../img/calendar3.png')

```
]
---

### Problem (pseudo formally)

* A series of $j \in \mathcal{J}$ missions are planned along a horizon divided into $t \in \mathcal{T}$ periods. Since all missions are already scheduled, we know time periods $T_j \subset \mathcal{T}$ in which they will be realized. 
* Each mission requires a certain number $r_{j}$ of resources $i \in \mathcal{I}$ which it employs for a time duration defined by $h_j$ in each period.
* Resources (heterogeneous) require recurrent preventive maintenance operations since the realization of missions diminish their remaining usage time. There are also min (max) distances between maintenances.
* A maintenance operation takes exactly $m$ periods and cannot be interrupted. It restores the resource's remaining usage time to exactly $H$ units. There is a limited capacity for maintenances.

--

* **Other**: soft constraints on the state of the fleet at each period and at the end, minimum assignments on missions, initial conditions, default consumption.

---

### Variables (subset)

* $a_{ijtt'}$ : has value one if aircraft $i$ starts an assignment to mission $j$ at the beginning of period $t$ and finishes at the end of period $t'$, zero otherwise.  
* $m_{ip}$ : has value one if aircraft $i$ uses pattern $p$, zero otherwise.
  * each pattern $p$ has a single feasible combination of maintenance starts for an aircraft during the whole planning (usually only two checks per aircraft).

--

.center[

```{r, echo=FALSE, out.width='100%'}
knitr::include_graphics('./../../../img/calendar3.png')

```
] 
---

## Prediction problem

---

### Distance between maintenances

.pull-left[
* The distance between maintenance has a maximum of 60 periods.
* Depending on the amount of missions to do in a planning horizon, this can be shorter.
* The more room we leave as possible distance between maintenances, the more patterns and, as a result, maintenance variables we have.

]

.pull-right[
```{r, echo=FALSE, out.width='100%'}
imgs = c('hist_mean_dist_complete_IT000125_20190716.png')
knitr::include_graphics(imgs)
```
]

---

### Forecasting + Optim

**We want to**:

1. Train a statistical model to predict the mean distance between maintenances for any given instance.
2. Use this information to limit the possible combinations of patterns to assign.

--

**Benefits**:

1. **Performance**: a smaller model is easier to solve.
2. **User feedback**: direct feedback about the solution without needing to solve any model.
3. **More stable solutions**: Every aircraft flies an amount that is closest to the mean of the fleet.

--

**The better we're able to predict the optimal distance between maintenances for the whole fleet, the less optimality we will lose**

---

### Implementation

* Done in python with scikit learn and statsmodels.
* Trained with 5000 small instances solved for an hour with CPLEX.
  * a 70/30 split was done in training/test.
  * solutions far from optimum were not used.
* Validated (semi) visually (not sure the good performance metric for predicting upper bounds).

--

I tested the following techniques:

* Linear regressions.
* Decision Tree regressions.
* Neural networks regressions.
* **Gradient Boosting regressions**.
* **Quantile regressions**.

--

The only two with support for quantile regression (i.e. bounds at quantiles 10% and 90%) were the last two. And the last one had the additional benefit of returning regression coefficients. So I chose that one.

---

### Model

.pull-left[
* **Technique**: *Quantile regressions* to estimate upper and lower bounds.
* **Input features**: 
  * mean flight demand per period (up to power 4), 
  * total remaining flight hours at start (init), 
  * variance of flight demand, 
  * demand of special missions, 
  * number of period where flight demand is cut in two.
* **Output features**: *mean distance between maintenances*.
]

--

.pull-right[
```{r, echo=FALSE, out.width='100%'}
imgs = c('QuantReg_mean_consum_upper_mean_dist_complete_IT000125_20190716.png')
knitr::include_graphics(imgs)
```
]

---

## Current results

---

### Experiments

* Number of instances: small (5000), medium (1000) and large (1000).
* Time limit at 3600 seconds.
* We seeded instance generation for better comparison.
* CPLEX running 1 thread.

Large instances have:

* 45 aircraft
* 90 months
* +20 missions (3 active missions at any given time).

---

### How good is it (performance)

Better solutions, more solutions.

.pull-left[
For instances were an integer solution was found:

* in ~85% of instances we outperformed base model.
* ~105% average gain in objective function.
* ~18% less time in average to solve.
]

--

.pull-right[
Many improvements in feasible solutions:

|indicator         | base| cuts| %|
|:-----------------|----:|----:|--------:|
|Infeasible        |   55|   71|    29.09|
|IntegerFeasible   |  183|  469|   156.28|
|IntegerInfeasible |  665|  291|   -56.24|
|Optimal           |   92|  166|    80.43|
|total_instances   |  995|  997|     0.20|
]
---

### How good is it (optimality)

For instances were an optimal solution was found (optimum degradation):
* 95% of instances had less than 4% gap with real optimal.

.center[
```{r, echo=FALSE, out.width='100%'}
imgs = c('relative_gap_optimality.png')
knitr::include_graphics(imgs)
```
*some outliers missing at right side...
]

---

## About my ambition.

---

### Present steps vs ideal steps

.pull-left[
What I have now:
```{r, out.width='100%'}
library(DiagrammeR)
mermaid("
graph TB
    i2==>|ML|s3
    subgraph Input data
    i1[raw input]-->|manual|i2[input features]
    end
    subgraph Solution
    s1[raw solution]-->s2[good patterns]
    s2-->|manual|s3[pattern features]
    s3==>|selection|s2
    end")
```
]

--

.pull-right[
What I want:
```{r, out.width='100%'}
mermaid("
graph TB
    i1==>|ML|s2
    subgraph Input data
    i1[raw input]-->|manual|i2[input features]
    end
    subgraph Solution
    s1[raw solution]-->s2[good patterns]
    s2-->|manual|s3[pattern features]
    end")
```
]

---

## Some questions

I want to be able to match assignment patterns (e.g. maintenances schedules) and unstructured input data in order to generalize a decomposition approach.

For that I need to know:

(from least ambitious to most ambitious):

* How do I train the other methods to predict bounds when they only measure mean squares (or some other function that returns the 'expected')?
* Is it possible to match input data to distance between maintenances without the need to manually generate the input features?
  * Assuming the amount of input data can, of course, change depending on the instance. I'm thinking in image recognition where images are all different size.
  * If so, how do I transform my input in such a way that can be read by some ML technique without having to obtain features? (e.g. do I need to create an image of it?)
* Is it possible to match input data to a set of patterns so that I don't even need to even tell the prediction model what characteristic of the patterns I want.