### 2020-02-14 R/presentations/roadef202002/presentation.Rmd
ROADEF2020 presentation. v1.

---
title: "An alternative MIP formulation for the Military Flight and Maintenance Planning problem"
author: |
    | Franco Peschiera, Alain Haït, Olga Battaïa, Nicolas Dupin
    | \tiny (franco.peschiera@isae-supaero.fr)
date: "February, 2020"
output:
  xaringan::moon_reader:
    css: ["footer-header.css", "default"]
    self_contained: false
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
    footer: "ISAE-SUPAERO, Université de Toulouse, France"
  slidy_presentation:
    css: ./../../../css/baobab_slidy.css
    incremental: yes
    footer: "ISAE-SUPAERO, Université de Toulouse, France"
    toc: true
  beamer_presentation:
    theme: "Berlin"
    toc: true
    slide_level: 2
    incremental: yes
    fig_height: 4
    includes:
        in_header: header.tex
editor_options: 
  chunk_output_type: console
---

layout: true

<div class="my-header"></div>

<div class="my-footer"><span>ISAE-SUPAERO, Université de Toulouse, France    
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;
ROADEF2020</span></div>

---

### Structure of talk

1. Problem.
1. State of the art.
1. MIP formulations.
2. Machine Learning.
2. Experiments.
3. Conclusions.

---

### Problem (informally)

Assign both **missions** and **maintenance operations** to a fleet of aircraft in order to maximize availability and minimize costs. **Missions** have fixed start and end times and have particular needs in terms of aircraft and flight hours. **Maintenances operations** have specific rules that govern their frequency and capacity.

--

.center[
```{r, echo=FALSE, out.width='100%', warning=FALSE}
library(RefManageR)
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
knitr::include_graphics('calendar3.png')

BibOptions(check.entries = FALSE,
           bib.style = "authoryear",
           cite.style = "alphabetic",
           style = "markdown",
           hyperlink = FALSE,
           dashed = FALSE)
myBib <- ReadBib("./MFMP.bib", check = FALSE)

```
]

---

### Problem

* A series of $j \in \mathcal{J}$ missions are planned along a horizon divided into $t \in \mathcal{T}$ periods. Since all missions are already scheduled, we know time periods $T_j \subset \mathcal{T}$ in which they will be realized. 
* Each mission requires a certain number $r_{j}$ of aircraft $i \in \mathcal{I}$ which it employs for $h_j$ hours in each period. An aircraft assigned to a mission $j$ must be assigned for at least $MT^{min}_{j}$ and at most $MT^{max}_{j}$ consecutive periods.
* The heterogeneous fleet requires recurrent preventive maintenance operations (checks) as a result of the total flight hours since the last check. There are also min (max) distances between maintenances.
* A check takes exactly $m$ periods and cannot be interrupted. There is a limited capacity for maintenances at each period.
* The objective is to minimize the total number of checks while maximizing the status of the fleet at the end of the horizon.

--

**Other**: soft constraints on the state of the fleet at each period, initial conditions, default consumption, reduce variance of frequency of checks.

---

### State of the art

* FMP: Flight and Maintenance Planning problem.
* In `r Citep(myBib, "Cho2011")`, US Army aircraft were assigned daily operations over a year to aircraft in order to minimize the maximum number of maintenances.
* In `r Citep(myBib, "Kozanidis2008")`, Greek aircraft had monthly assignments of maintenances and flight hours in order to maximize the availability and final state of squadrons.
* In `r Citep(myBib, "Verhoeff2015")`, monthly assignments were done and several objectives were taken into account: availability, serviceability and final state.
* In `r Citep(myBib, "Seif2018")`, a generalization for different types of maintenances and capacities was done.
* In `r Citep(myBib, "Peschiera2019")`, the problem was proved NP-Complete.

---

### Previous formulation

* $a_{jti}$   :  =1 if mission $j \in J$ in period $t \in \mathcal{T}_j$ is realized with aircraft $i \in \mathcal{I}_j$, 0 otherwise.  
* $a^s_{jti}$ :  =1 if aircraft $i$ starts a new assignment to mission $j$ in period $t$. If $a_{jti} = 1$ and $a_{j(t-1)i} = 0$.  
* $m_{it}$    :  =1 if aircraft $i \in I$ starts a check in period $t \in \mathcal{T}$, 0 otherwise.  
* $u_{it}$    :  flown time (continuous) by aircraft $i \in I$ during period $t \in \mathcal{T}$.

\begin{align}
 & u_{it} \geq \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j 
    & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:flight1}\\
 & u_{it} \geq U^{min} (1 - \sum_{t' \in \mathcal{T}^s_t} m_{it'})
    & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:flight2}\\
 & rft_{it} \leq rft_{i(t-1)} + H^M m_{it} - u_{it}
    & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:first_rft_upper}\\
& rft_{it} \in [0,H^M]
      & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:first-mu}
\end{align}

---

### New formulation

* $a_{ijtt'}$ : =1 if aircraft $i$ starts an assignment to mission $j$ at the beginning of period $t$ and finishes at the end of period $t'$, zero otherwise.  
* $m_{ip}$ : =1 if aircraft $i$ uses check pattern $p$, zero otherwise.
  * each pattern $p$ has a single feasible combination of check starts for an aircraft during the whole planning (usually only 1-2 checks per aircraft).

\begin{align}
& \sum_{(j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{ic}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{tc} \leq H^{max} + H^{max} (1 - m_{ip}) & \notag \\
  & \hspace{200px}  i \in \mathcal{I}, p \in \mathcal{P}, c \in \mathcal{C}_p \label{eq:cycle_hours2}\\
\end{align}

---

### Formulation

\begin{align}
  & \text{Max}\;
  \sum_{i \in \mathcal{I}, p \in \mathcal{P}} m_{ip} \times W_p \\
  & \sum_{i \in \mathcal{I}, p \in \mathcal{P}_{t}} m_{ip} \leq C^{max} 
          & t \in \mathcal{T} \label{eq:capacity1}\\
  & \sum_{i \in \mathcal{I}_j, (t_1, t_2) \in \mathcal{T}_{jt}} a_{ijt_1t_2} \geq R_j
          & j \in \mathcal{J}, t \in \mathcal{TJ}_j  \label{eq:missionres}\\
  & \sum_{p \in \mathcal{P}_{t}} m_{ip} + \sum_{j \in \mathcal{J}_t \cap \mathcal{J}_i} \sum_{(t_1, t_2) \in \mathcal{T}_{jt}} a_{ijt_1t_2} \leq 1 
          & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}\\
  & \sum_{(j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{ic}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{tc} \leq H^{max} + H^{max} (1 - m_{ip}) & \notag \\
    & & i \in \mathcal{I}, p \in \mathcal{P}, c \in \mathcal{C}_p \\
\end{align}

---

### Distance between maintenances

.pull-left[
* The distance between maintenance has a maximum of $E^{max}$ periods.
* Depending on the instance, the optimal distance can be shorter.
* This distance conditions the total number of patterns to create.
]

.pull-right[
```{r, echo=FALSE, out.width='100%'}
imgs = c('hist_mean_dist_complete_IT000125_20190716.png')
knitr::include_graphics(imgs)
```
]

---

### Forecasting + Optim

**We want to**:

1. Train a statistical model to predict the mean distance between maintenances for any given instance.
2. Use this information to limit all possible combinations of patterns to generate.

--

**Benefits**:

1. **Performance**: a smaller model is easier to solve.
2. **User feedback**: direct feedback about the solution without needing to solve any model.
3. **More stable solutions**: Every aircraft flies an amount that is closest to the mean of the fleet.

--

**The better we're able to predict the optimal distance between maintenances for the whole fleet, the less optimality we will lose**

---

### Prediction model

.pull-left[
* **Technique**: *Quantile regressions* to estimate upper and lower bounds.
* **Training**: 5000 small instances.
* **Input features**: 
  * mean flight demand per period,
  * total remaining flight hours at start (init), 
  * variance of flight demand, 
  * demand of special missions, 
  * number of period where flight demand is cut in two.
* **Output features**: mean distance between maintenances.
]

--

.pull-right[
```{r, echo=FALSE, out.width='100%'}
imgs = c('QuantReg_mean_consum_upper_mean_dist_complete_IT000125_20190716.png')
knitr::include_graphics(imgs)
```
]

---

### Experiments

* Number of instances: medium (1000), large (1000) and very large (1000).
* Time limit at 3600 seconds.
* We seeded instance generation for better comparison.
* CPLEX running 1 thread.

Largest instances have 60 aircraft, 90 periods, ~30 missions (4 active missions at any given time).

1. Create forecasting model based in 5000 small instances.
2. Use forecasting model to predict bounds on distance between maintenaances: $\hat{\mu}_{t'-t}^{lb}$,  $\hat{\mu}_{t'-t}^{ub}$.
3. Implement the pseudo-cut:

\begin{align}
    & m_{ip} = 0 & p_{t'} - p_t < \hat{\mu}_{t'-t}^{lb} - tol \label{eq:dist_lb} \\
    & m_{ip} = 0  &  p_{t'} - p_t > \hat{\mu}_{t'-t}^{ub} + tol \label{eq:dist_ub}
\end{align}

---

### How good is it (performance)

Faster solutions, more solutions.

--

.center[
```{r, echo=FALSE, out.width='80%'}
imgs = c('time_performance_ordered_2tasks.png')
knitr::include_graphics(imgs)
```
]

---

### How good is it (optimality)

For instances were an optimal solution was found (optimum degradation):
* 95% of instances had less than 4% gap with real optimal.

.center[
```{r, echo=FALSE, out.width='80%'}
imgs = c('quality_degradation_2tasks.png')
knitr::include_graphics(imgs)
```
]

---

## Further steps

* **Predict a distribution** and sample patterns from the distribution instead of predicting patterns.
* **Warm-start Column Generation** with a selected subset of potentially good patterns.
* **Automatize prediction** so it can be easily integrated in other problems.

---

# References

```{r refs, echo=FALSE, results="asis"}
PrintBibliography(myBib)
```