### 2020-09-04 R/presentations/EDAA202003/presentation.Rmd
changes to EDAA202003 ppt.

---
title: "Predicting solutions for the Military Flight and Maintenance Planning problem"
author: |
    | Franco Peschiera, Alain Haït, Olga Battaïa, Nicolas Dupin
    | \tiny (franco.peschiera@isae-supaero.fr)
date: "September, 2020"
output:
  xaringan::moon_reader:
    css: ["footer-header.css", "default"]
    self_contained: false
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
    footer: "ISAE-SUPAERO, Université de Toulouse, France"
  slidy_presentation:
    css: ./../../../css/baobab_slidy.css
    incremental: yes
    footer: "ISAE-SUPAERO, Université de Toulouse, France"
    toc: true
  beamer_presentation:
    theme: "Berlin"
    toc: true
    slide_level: 2
    incremental: yes
    fig_height: 4
    includes:
        in_header: header.tex
editor_options: 
  chunk_output_type: console
---

```{r, echo=FALSE, warning=FALSE}
library(RefManageR)
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)

BibOptions(check.entries = FALSE,
           bib.style = "authoryear",
           cite.style = "alphabetic",
           style = "markdown",
           hyperlink = FALSE,
           dashed = FALSE)
myBib <- ReadBib("./MFMP.bib", check = FALSE)

```

layout: true

<div class="my-header"></div>

<div class="my-footer"><span>ISAE-SUPAERO, Université de Toulouse, France    
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;
EDAA Journée doctorants 2020</span></div>

---

### Structure of talk

1. Problem.
2. State of the art.
1. Solving techniques.
2. Machine Learning.
2. Experiments.
3. Conclusions.

---

### Problem (informally)

Assign both **missions** and **maintenance operations** to a fleet of aircraft in order to maximize availability and minimize costs. **Missions** have fixed start and end times and have particular needs in terms of aircraft and flight hours. **Maintenance operations** have specific rules that govern their frequency and capacity.

--

.center[
```{r, echo=FALSE, out.width='100%', warning=FALSE}
knitr::include_graphics('calendar3.png')

```
]

---

### Military aircraft maintenance

.pull-left[

* Mirage 2000 fleet.
* A single workshop to do the maintenance operations.
* Check type D (overhaul maintenances): 
  * every 5 years or 1000 flight hours and 
  * takes 4-6 months.
* Fleets of 15-60 aircraft.
* Planning horizons of 90 periods (= 7.5 years).

]

.pull-right[
```{r, echo=FALSE, out.width='100%', warning=FALSE}
knitr::include_graphics('Mirage_2000C.jpg')
```

*French Mirage 2000C. Public Domain*
]

---

### Problem

* A series of $j \in \mathcal{J}$ pre-scheduled missions are planned along a horizon divided into $t \in \mathcal{T}$ periods.
* Each mission requires a certain **number** $R_{j}$ of aircraft $i \in \mathcal{I}$ which it employs for $H_j$ **hours** in each period.
* The heterogeneous fleet requires recurrent preventive maintenance operations (**checks**).
* A check takes exactly $M^d$ periods and cannot be interrupted. There is a limited **capacity** for maintenances at each period.
* The objective is to **minimize the total number of checks** while **maximizing the status of the fleet** at the each period of the horizon.

--

**Other**: soft constraints on the state of the fleet at each period, initial conditions, default consumption, reduce variance of frequency of checks.

---

### State of the art

* FMP: Flight and Maintenance Planning problem.
* In `r Citep(myBib, "Cho2011")`, US Army aircraft were assigned daily operations over a year to aircraft in order to minimize the maximum number of maintenances.
* In `r Citep(myBib, "Kozanidis2008")`, Greek aircraft had monthly assignments of maintenances and flight hours in order to maximize the availability and final state of squadrons.
* In `r Citep(myBib, "Verhoeff2015")`, monthly assignments were done and several objectives were taken into account: availability, serviceability and final state.
* In `r Citep(myBib, "Seif2018")`, a generalization for different types of maintenances and capacities was done.
* In `r Citep(myBib, "Peschiera2019")`, the problem was proved NP-Complete.

---

### "Classical" solving techniques

$a_{ijt_1t_2}$: mission $j$ is assigned to aircraft $i$ between $t_1$ and $t_2$.  
$m_{it_1t_2}$: aircraft $i$ starts a check in $t_1$ and another check in $t_2$.

\begin{align}
  & \text{Max}\;
  \sum_{i \in \mathcal{I}, p \in \mathcal{P}} m_{ip} \times W_p \\
  & \sum_{i \in \mathcal{I}, p \in \mathcal{P}_{t}} m_{ip} \leq C^{max} 
          & t \in \mathcal{T} \label{eq:capacity1}\\
  & \sum_{i \in \mathcal{I}_j, (t_1, t_2) \in \mathcal{T}_{jt}} a_{ijt_1t_2} \geq R_j
          & j \in \mathcal{J}, t \in \mathcal{TJ}_j  \label{eq:missionres}\\
  & \sum_{p \in \mathcal{P}_{t}} m_{ip} + \sum_{j \in \mathcal{J}_t \cap \mathcal{J}_i} \sum_{(t_1, t_2) \in \mathcal{T}_{jt}} a_{ijt_1t_2} \leq 1 
          & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}\\
  & \sum_{(j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{ic}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{tc} \leq H^M + M (1 - m_{ip}) & \notag \\
    & & i \in \mathcal{I}, p \in \mathcal{P}, c \in \mathcal{C}_p \\
\end{align}

---

### Forecasting + Optimization

**We want to**:

1. Train a statistical model to predict characteristics of good (optimal?) solutions for any given instance.
2. Use this information to limit the solution space and make it easier to find those potentially good solutions.

--

**Benefits**:

1. **Performance**: a smaller problem is easier to solve.
2. **User feedback**: direct feedback about the solution without needing to solve any model.
3. **More stable solutions**: Every aircraft flies an amount that is closest to the mean of the fleet.

--

**The better we're able to predict those characteristics for the whole fleet, the less optimality we will lose**

---

### Prediction model

* **Technique**: *Quantile regressions* to estimate upper and lower bounds.
* **Training**: 5000 small instances.
* **Input features**: 
  * mean flight demand per period,
  * total remaining flight hours at start (init), 
  * variance of flight demand, 
  * demand of special missions, 
  * number of period where flight demand is cut in two.
* **Output features**: mean distance between maintenances.

--

```{r, echo=FALSE, out.width='100%', warning=FALSE}
knitr::include_graphics('calendar3.png')
```

---

### Experiments

* Number of instances: medium (1000), large (1000) and very large (1000).
* Time limit at 3600 seconds.
* We seeded instance generation for better comparison.
* CPLEX running 1 thread.

--

Largest instances have 60 aircraft, 90 periods, ~30 missions (4 active missions at any given time).

1. Create forecasting model based in 5000 small instances.
2. Use forecasting model to predict bounds on distance between maintenances: $\hat{\mu}_{t'-t}^{lb}$,  $\hat{\mu}_{t'-t}^{ub}$.
3. Implement the pseudo-cut:

\begin{align}
    & m_{ip} = 0 & p_{t'} - p_t < \hat{\mu}_{t'-t}^{lb} - tol \label{eq:dist_lb} \\
    & m_{ip} = 0  &  p_{t'} - p_t > \hat{\mu}_{t'-t}^{ub} + tol \label{eq:dist_ub}
\end{align}

4. Recycling.

---

### How good is it (performance)

Faster solutions, more solutions.

--

.center[
```{r, echo=FALSE, out.width='80%'}
imgs = c('time_performance_ordered_2tasks.png')
knitr::include_graphics(imgs)
```
]

---

### How good is it (optimality)

For instances were an optimal solution was found (optimum degradation):
* 95% of instances had less than 4% gap with real optimal.

.center[
```{r, echo=FALSE, out.width='80%'}
imgs = c('quality_degradation_2tasks.png')
knitr::include_graphics(imgs)
```
]

---

## Further steps

* **Better predictions** with better features, or predicting several characteristics of optimal solutions.
* **Predict a distribution** and sample patterns from the distribution instead of predicting patterns.
* **Warm-start Column Generation** with a selected subset of potentially good patterns.
* **Automatize prediction** so it can be easily integrated in other problems.

---

# References

```{r refs, echo=FALSE, results="asis"}
PrintBibliography(myBib)
```